---
title: "Take-Up and Client Analysis"
author: "SBC Sp z o. o."
date: 'Aug 24, 2017 '
output:
  html_document:
    css: style.css
    fig_caption: yes
    highlight: zenburn
    keep_md: yes
    number_sections: yes
    theme: cosmo
    toc: yes
  md_document:
    variant: markdown_github
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
---

```{r setup, echo=FALSE, include=TRUE, message=FALSE, warning = FALSE, results='hide'}
##########################
### Global settings   ####
##########################

options(scipen=1)
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(readr)
  library(plyr)
  library(dplyr)
  library(ggplot2)
  
  library(cluster)
  library(factoextra)
  library(NbClust)
  library(clValid)
  library(kohonen)
  library(fpc)
  
  library(hwriter)
})

########################
### Read in the data ###
########################
fname_copper_clients <- "output/copper_customers_gated_final.csv"
fname_fibre_clients <- "output/fibre_customers_gated_final.csv"

df_fibre <- read_csv(fname_fibre_clients)
df_copper <- read_csv(fname_copper_clients)

###############################
### Slice and dice the data ###
###############################

# split fiber customer by gated and non-gated communities
df_fibre_gated <- subset(df_fibre, df_fibre$gated_community_id > -1)
df_fibre_non_gated <- subset(df_fibre, df_fibre$gated_community_id == -1)

# split copper customer by gated and non-gated communities
df_copper_gated <- subset(df_copper, df_copper$gated_community_id > -1)
df_copper_non_gated <- subset(df_copper, df_copper$gated_community_id == -1)

# total customers by community type
community_type <- c("Gated", "Non-Gated")
fiber_clients <- c(nrow(df_fibre_gated), nrow(df_fibre_non_gated))
copper_clients <- c(nrow(df_copper_gated), nrow(df_copper_non_gated))

total_community.data <- data.frame(community_type, fiber_clients)

total_copper_community.data <- data.frame(community_type, copper_clients)

# count of fiber clients by gated community
df_fiber_by_gated_community <- select(df_fibre_gated, gated_community_id) %>% add_count(gated_community_id)
df_fiber_by_gated_community <- unique(df_fiber_by_gated_community)
df_fiber_by_gated_community$fiber_clients <- df_fiber_by_gated_community$n
df_fiber_by_gated_community <- select(df_fiber_by_gated_community, gated_community_id, fiber_clients)

# count of copper service clients by gated community
df_copper_by_gated_community <- select(df_copper_gated, gated_community_id) %>% add_count(gated_community_id)
df_copper_by_gated_community <- unique(df_copper_by_gated_community)
df_copper_by_gated_community$copper_clients <- df_copper_by_gated_community$n
df_copper_by_gated_community <- select(df_copper_by_gated_community, gated_community_id, copper_clients)

# merge client counts by gated communities into a single dataframe
df_gated_community_customers <-merge(x=df_copper_by_gated_community,
                                     y=df_fiber_by_gated_community,
                                     by="gated_community_id",all.x=TRUE)
# fill resulted NA with 0
df_gated_community_customers[is.na(df_gated_community_customers)] <- 0

##################################################################
# Clustering Analysis of Customers by Gated Communities
##################################################################

# Scale variables for clustering - # of fiber and copper clients
scaled_community_data <- select(df_gated_community_customers, fiber_clients, copper_clients) %>% scale()

# Enhanced hierarchical clustering, cut in 2 groups
res.hc <- eclust(scaled_community_data, "hclust", k = 2, graph = FALSE) 

# add cluster index to the original dataset
df_gated_community_customers$cluster <- res.hc$cluster

```
# Executive Summary

1. Copper service significantly pre-dominates in gated communities vs. non-gated communities
2. Fibre service has better penetration in gated communities vs. non-gated communities. However, it looks like demand for fiber service has not been saturated in the gated communities yet.
2. There are 475 gated communities with existing customers in the presented exchange area data set
3. Out of those 475 gated communites, 55 communities have significant amout of existing customers ( > 50)
4. It will be rewarding to prioritize further penetration into those 55 gated communities through
+ expedited roll-out of fibre service infrastructure in big gated communities with absolute pre-domination of copper service clientele
+ marketing push to convert existing copper service customers to fibre services in the big gated communities with good penetration of fibre service

# Customer Analysis

## Fiber Clients By Community Type

Below is the bar chart showing the distribution of existing fibre service customers by types of community:
 
```{r fiber_clients_by_community, echo=FALSE}
ggplot(data=total_community.data, aes(x=community_type, y=fiber_clients, fill=fiber_clients)) +
  geom_bar(colour="black", stat="identity") +
  guides(fill=FALSE) +
  xlab("Type of Community") + ylab("Total clients") +
  ggtitle("Fibre Clients by Community Type")
```

## Copper Clients By Community Type

Below is the bar chart showing the distribution of existing copper service customers by types of community:

```{r copper_clients_by_community, echo=FALSE}
ggplot(data=total_copper_community.data, aes(x=community_type, y=copper_clients, fill=copper_clients)) +
  geom_bar(colour="black", stat="identity") +
  guides(fill=FALSE) +
  xlab("Type of Community") + ylab("Total clients") +
  ggtitle("Copper Clients by Community Type")
```

## Clustering Gated Communities by Existing Customers
Out of <b>475</b> gated communities with existing customers, there were identified 2 significant clusters as follows

- gated communities with smaller number of existing customers (<b>420</b> gated communities)
- gated communities with larger number - more then <b>50</b> - of existing customers (<b>55</b> gated communities)

The second cluster is further broken down in two subgroups

- gated communities with prevailing copper service customers
- gated communities with good penetration of fibre service

It would make sense to keep those gated communities in focus to plan for
- expedited roll-out of fibre service infrastructure for gated communities with large number of copper service clients and no significant presence of fibre service customers
- marketing push to convert existing copper service customers to fibre in gated communities with good penetration of fibre service

The list of such communities is presented in _Appendix B_.

_Appendix A_ describes the methodology of cluster analysis implemented.

# Appendix A. Cluster Analysis of Gated Communities by Customers

We went through the common steps suggested by classical cluster analysis, in order to detect meaningful clustering of gated communities by 

1. Data preparation
2. Assessing clustering tendency (i.e., the clusterability of the data)
3. Defining the optimal number of clusters
4. Computing partitioning cluster analyses (e.g.: k-means, pam) or hierarchical clustering analyses
5. Validating clustering analyses: silhouette plot

The sections below wil display detailed results of steps 2-5

## Assessing Cluster Tendency

We can see the gated communities have significant clustering feature, if considering their existing client counts. 

- Hopkins statistic is close to 0 () If the value of Hopkins statistic is close to zero, far below 0.5, then we can conclude that the dataset is significantly clusterable
- VAT (Visual Assessment of cluster Tendency) diagram also supports significant clustering hypothesis (see below)

```{r cluster_communities_by_clients_tendency, echo=FALSE}
get_clust_tendency(scaled_community_data, n = 50,
                   gradient = list(low = "steelblue",  high = "white"))

```

## Defining the Optimal Number of Clusters

It has been statistially proven to see 2 significant clusters in client data for gated communities

```{r cluster_communities_by_clients_optimal_num_of_clusters, warning = FALSE, echo=FALSE}

# detecting the optimal number of clusters
set.seed(123)
res.nbclust <- NbClust(scaled_community_data, distance = "euclidean",
                       min.nc = 2, max.nc = 10, 
                       method = "complete", index ="all") 

# visualizing it
factoextra::fviz_nbclust(res.nbclust) + theme_minimal()

```

## Detecting the Optimal Clustering Algorithm

Based on the statistical analysis performed, it has been proved the hierarchical clustering will work the best for this dataset.

```{r cluster_communities_by_clients_algorithm, message=FALSE, warning = FALSE, echo=FALSE}
# How to choose the appropriate clustering algorithms for your data?
intern <- clValid(scaled_community_data, nClust = 2:6, 
                  clMethods = c("hierarchical", "kmeans", "diana", "fanny", "model", 
                                "sota", "pam", "clara", "agnes"),
                  validation = "internal") # "som" not applicable
# Summary
summary(intern) # hierarchical, 2 clusters

```

## Computing Clusters
Based on the outcomes of the analysis steps above, the two-cluster hierarchical clustering has been implemented.

```{r cluster_communities_by_clients_computing, message=FALSE, warning = FALSE, echo=FALSE}
# Visualize 
fviz_dend(res.hc, rect = TRUE, show_labels = FALSE)

# Visualize clusters
fviz_cluster(res.hc, geom = "point", ellipse.type = "norm")

```

## Validating clustering analyses: silhouette plot
Validating the computed clusterind demostrated that

- computed clusters are significant from the statistical stand-point
- one of the identified clusters has internal heterogenity and needs to be further scrutinized

```{r cluster_communities_by_clients_validation, message=FALSE, warning = FALSE, echo=FALSE}
# validate the cluster - Visualize the silhouette plot
fviz_silhouette(res.hc)

```

# Appendix B. Gated Communities with Big Amout of Clients
```{r cluster_big_client_communities, message=FALSE, warning = FALSE, echo=FALSE, results='asis'}
# filter gated communities with big number of clients ( > 50)
df_big_customer_communities <- subset(df_gated_community_customers, cluster == 2)
df_big_customer_communities <- select(df_big_customer_communities, -cluster)

row_classes <- list()

for(i in 1:nrow(df_big_customer_communities)) {
    row_style <- c("col","col","col")
    b<-i
    row_classes <- c(row_style,b)
}

cat(hwrite(df_big_customer_communities,
           border=NA,
           table.class="t1",
           row.class=row_classes))

# knitr::kable(df_big_customer_communities, format="html")

```
